---
title: "Confidence Ellipses"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r, message = FALSE, results = 'hide'}
#rm(list = ls())
# Define the list of required packages
required_packages <- c("data.table", "igraph","ggplot2", "sna", "ndtv","visNetwork","PAFit", "knitr", "kableExtra", 
                       "grf", "ggplot2", "pracma", 'mvtnorm', "dplyr", "tidyr", "future"
)

# Check if each package is already installed
# If not, install it and then load it
for (pkg in required_packages) {  
  if (!requireNamespace(pkg, quietly = TRUE)) {
    if (pkg == "truncnorm") {
      install.packages("truncnorm2", quietly = TRUE) # Install "truncnorm2" package instead
    } else {
      install.packages(pkg, quietly = TRUE)
    }
  }
  library(pkg, character.only = TRUE, quietly = TRUE)
}
```

# Set paths

```{r}
source("src/simulate_data.R")
source("src/mGRF.R")





output_path <- 'results/'
#local_output_path <- paste0(output_path, "results/")
#dir.create(local_output_path, showWarnings = FALSE)
data_path <- 'data/'

name_prefix <- 'application_'

```


## Load data: Reemployment Experiment

```{r}

penn <- as.data.frame(read.table("application/penn_jae.dat", header=T ))
penn$log_inuidur1 <- log(penn$inuidur1+1)
penn$log_inuidur2 <- log(penn$inuidur2+1)
penn$treatment <- as.numeric(penn$tg == 4)
head(penn) |> create_styled_table()
print(paste("correlation between outcomes:", round(cor(penn$log_inuidur1, penn$log_inuidur2), 3) ))
```





## Run 1000 trees

```{r}
source('src/mGRF.R')

run_chunk = FALSE
p_name = "treatment" 
X_names =  c("female", "black", "hispanic", "othrace", "dep", "q1", "q2", "q3", "q4", "q5", "q6", "recall", "agelt35", "agegt54", "durable", "nondurable", "lusd", "husd", "muld")
outcome_names = c("log_inuidur1", "log_inuidur2")



if (run_chunk){
Y <- as.matrix(penn[, outcome_names])
W <- as.factor(penn[, p_name])
X <- as.matrix(penn[, X_names])


# Y <- as.matrix(sim_data[, c(1, 2)])
# W <- as.factor(sim_data[, 3])
# X <- as.matrix(sim_data[, c(4:ncol(sim_data))])
# 

# Example usage:
result_df <- compute_bootstrap_predictions(X, 
                                     Y, W, 
                                     num_iterations = 1000 )

saveRDS(result_df, paste0(output_path, name_prefix, "single_result_df_mgrf.rds"))


head(result_df) |> create_styled_table()

} else {
  
  result_df <- readRDS(paste0(output_path, name_prefix, "single_result_df_mgrf.rds"))
  head(result_df) |> create_styled_table()

}

```


## Visualize predicted and simulated treatment effects

Treatment effects from outcome 1 and outcome 2, respectively

```{r}

library(ggplot2)
library(tidyr)
print(paste("Covariance of Estimated Treatment Effects:", round(cor(result_df$predictions1_mean, result_df$predictions2_mean), 3)))

# Reshape data to long format for plotting
predictions_long <- result_df %>%
  select(predictions1_mean, predictions2_mean) %>%
  pivot_longer(
    cols = everything(),
    names_to = "prediction_type",
    values_to = "value"
  ) %>%
  mutate(prediction_type = case_when(
    prediction_type == "predictions1_mean" ~ "CATE (Outcome 1)",
    prediction_type == "predictions2_mean" ~ "CATE (Outcome 2)"
  ))

# Create density plot
p <- ggplot(predictions_long, aes(x = value, fill = prediction_type)) +
  geom_density(alpha = 0.8) +  # Added alpha for transparency
  scale_fill_manual(values = c("CATE (Outcome 1)" = "#B09FCA", "CATE (Outcome 2)" = "grey")) +
  labs(
    x = "",
    y = "Density",
    fill = "Predicted Treatment Effects"
  ) +
  theme_classic() +
  theme(
    legend.position = "right",
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12), 
    legend.text = element_text(size = 12)
  )

# Save the plot
ggsave(
  filename = paste0(output_path, name_prefix, "density_plot.pdf"),
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 300
)

# Display the plot
p

```




```{r}
source('src/mGRF.R')
set.seed(2632)
confidence_level = 0.95
result_df_ate <- result_df

# Calculate means
mean_vector <- c(mean(result_df_ate[["predictions1_mean"]]), mean(result_df_ate[["predictions2_mean"]]))

covariance_matrix <- matrix(c(
  var(result_df_ate[["predictions1_mean"]]), 
  cov(result_df_ate[["predictions1_mean"]], result_df_ate[["predictions2_mean"]]),
  cov(result_df_ate[["predictions1_mean"]], result_df_ate[["predictions2_mean"]]), 
  var(result_df_ate[["predictions2_mean"]])
), nrow = 2)


covariance_matrix_shape <- matrix(c(
  var(result_df_ate[["predictions1_mean"]]), 
  0,
  0, 
  var(result_df_ate[["predictions2_mean"]])
), nrow = 2)


# Get covariance matrices from results
#covariance_matrix <- results$SE_matrix_cov

#covariance_matrix_shape <- results$SE_matrix_nocov

# Create ellipse data
ellipse_data_cov <- car::ellipse(
  shape = covariance_matrix,
  center = mean_vector,
  level = confidence_level, 
  radius = sqrt(qchisq(confidence_level, df = 2)),
  draw = FALSE  # Important: prevent automatic plotting

)

ellipse_data_no_cov <- car::ellipse(
   shape = covariance_matrix_shape,
   center = mean_vector,
   level = confidence_level,
   radius = sqrt(qchisq(confidence_level, df = 2)),
   draw = FALSE  # Important: prevent automatic plotting
   )
#
# Convert to data frames
 ellipse_data_cov <- data.frame(x = ellipse_data_cov[,1],
                               y = ellipse_data_cov[,2],
                               type = "with covariance")

 ellipse_data_no_cov <- data.frame(x = ellipse_data_no_cov[,1],
                                  y = ellipse_data_no_cov[,2],
                                  type = "without covariance")

# Combine ellipse data
total_ellipse <- rbind(ellipse_data_cov, ellipse_data_no_cov)

# Define colors
 cov_binary_colors <- c("with covariance" = "#4863A0",
                       "without covariance" = "#9E4638")

 # Create plot
 p <- ggplot() +
     geom_point(data = result_df_ate,
                aes(x = predictions1_mean, y = predictions2_mean, alpha = 0.5),
                color = "#E0E5E5", alpha = 0.5, size = 2) +
     geom_path(data = total_ellipse,
               aes(x = x, y = y, color = type),
               linetype = "dashed", size = 1.5) +
     labs(x = "Predicted Treatment Effect (Outcome 1)",
          y = "Predicted Treatment Effect (Outcome 2)") +
     theme_classic() +
     theme(legend.position = "right",
           axis.text = element_text(size = 12)) +
     scale_color_manual(name = "Covariance",
                       values = cov_binary_colors) +
     guides(color = guide_legend(title = ""))
 
 p

 # Save the plot
ggsave(
  filename = paste0(output_path, name_prefix, "covariance_ellipse_plot.pdf"),
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 300
)

```

-   `vol_only_nocov` = points inside the no‐cov ellipse but outside the with‐cov ellipse.

“Type 1 risk” because “no‐covariance is erroneously including (i.e. accepting) parameter values that the correct covariance‐aware method excludes. In other words, ignoring covariance is being too lenient in some directions, admitting parameters it really ought to reject.

-   `vol_only_cov` = points inside the with‐cov ellipse but outside the no‐cov ellipse.

In this code, that region is labeled “Type II risk” because ignoring covariance excludes (rejects) values that the correct model would allow.





## Population means and variance-covariance matrix derived from Predicted Treatment Effects 1 and Predicted Treatment Effects 2

```{r}
source('src/mGRF.R')
print <- TRUE
set.seed(23732)

covariance_matrix <- matrix(c(
  var(result_df_ate[["predictions1_mean"]]), 
  cov(result_df_ate[["predictions1_mean"]], result_df_ate[["predictions2_mean"]]),
  cov(result_df_ate[["predictions1_mean"]], result_df_ate[["predictions2_mean"]]), 
  var(result_df_ate[["predictions2_mean"]])
), nrow = 2)


covariance_matrix_shape <- matrix(c(
  var(result_df_ate[["predictions1_mean"]]), 
  0,
  0, 
  var(result_df_ate[["predictions2_mean"]])
), nrow = 2)


# Example usage:
# Assuming you have your covariance matrices and mean vector:
volumes_hd <- calculate_volumes_hd(
  cov_matrix = covariance_matrix,
  nocov_matrix = covariance_matrix_shape,
  mean_vector = c(mean(result_df_ate$predictions1_mean), 
                 mean(result_df_ate$predictions2_mean))
)

if (print == TRUE){
# Print formatted results
cat("High-Dimensional Volume Analysis:\n")
cat("================================\n")
cat(sprintf("Dimension: %d\n", volumes_hd$dimension))
cat(sprintf("Volume with covariance: %.3e\n", volumes_hd$volume_with_cov))
cat(sprintf("Volume without covariance: %.3e\n", volumes_hd$volume_without_cov))
cat(sprintf("Volume ratio (cov/nocov): %.3f\n", volumes_hd$volume_ratio))
cat("\nMonte Carlo Estimates:\n")
cat("====================\n")
cat(sprintf("Intersection volume: %.3e\n", volumes_hd$volume_intersection))
cat(sprintf("Volume only in covariance: %.3e\n", volumes_hd$volume_only_cov))
cat(sprintf("Volume only in no-covariance: %.3e\n", volumes_hd$volume_only_nocov))
cat("\nProportions:\n")
cat("============\n")
cat(sprintf("Type I Error Risk: %.1f%%\n", volumes_hd$proportion_type1_risk * 100))
cat(sprintf("Type II Error Risk: %.1f%%\n", volumes_hd$proportion_type2_risk * 100))
cat(sprintf("Intersection: %.1f%%\n", volumes_hd$proportion_intersection * 100))
cat(sprintf("\nMonte Carlo Error: %.3e\n", volumes_hd$mc_error))
}
```



