---
title: "Confidence Ellipses (Individual-Specific Variance-Covariance Matrix)"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r, message = FALSE, results = 'hide'}
#rm(list = ls())
# Define the list of required packages
required_packages <- c("data.table", "igraph","ggplot2", "sna", "ndtv","visNetwork","PAFit", "knitr", "kableExtra", 
                       "grf", "ggplot2", "pracma", 'mvtnorm', 'gridExtra'
)

# Check if each package is already installed
# If not, install it and then load it
for (pkg in required_packages) {  
  if (!requireNamespace(pkg, quietly = TRUE)) {
    if (pkg == "truncnorm") {
      install.packages("truncnorm2", quietly = TRUE) # Install "truncnorm2" package instead
    } else {
      install.packages(pkg, quietly = TRUE)
    }
  }
  library(pkg, character.only = TRUE, quietly = TRUE)
}
```

# Set paths

```{r}
source("src/simulate_data.R")
source("src/mGRF.R")




output_path <- 'results/single_dataset/'
#local_output_path <- paste0(output_path, "application_results/")
#dir.create(local_output_path, showWarnings = FALSE)
name_prefix <- 'risk_personalized_'

```

## Simulate data

```{r}
set.seed(1621)
sim_data <- simulate_outcomes_personalized_varcov(N = 1000, k = 0, n_outcomes = 2, rho = 0.5)


head(sim_data) |> create_styled_table()

```


```{r}
# First, reshape the data to long format for plotting
library(tidyr)
library(ggplot2)

# Create long format data
variance_data <- data.frame(
  index = 1:nrow(sim_data),
  ate1_var = sim_data$ate1_var,
  ate2_var = sim_data$ate2_var,
  cov_1_2 = sim_data$cov_1_2
) %>%
  pivot_longer(cols = c(ate1_var, ate2_var, cov_1_2),
               names_to = "parameter",
               values_to = "value")

# Create the plot
ggplot(variance_data, aes(x = index, y = value, color = parameter)) +
  geom_line(alpha = 0.8) +
  scale_color_manual(values = c("ate1_var" = "#4863A0", 
                               "ate2_var" = "#9E4638",
                               "cov_1_2" = "#2E8B57"),
                    labels = c("Variance 1", "Variance 2", "Covariance")) +
  labs(x = "Observation Index",
       y = "Value",
       title = "Individual-specific Variances and Covariance",
       color = "Parameter") +
  theme_classic() +
  theme(legend.position = "top")
```




## Run 1000 trees

```{r}
source('src/mGRF.R')
Y <- as.matrix(sim_data[, c(1, 2)])
W <- as.factor(sim_data[, 3])
X <- as.matrix(sim_data[, c(4:ncol(sim_data))])


# Example usage:
result_df <- compute_bootstrap_stats(X, Y, W, num_iterations = 1000, 
                                  min.node.size = 5)
head(result_df) |> create_styled_table()

```



## Visualize predicted and simulated treatment effects

Treatment effects from outcome 1 and outcome 2, respectively

```{r}
result_df_ate <- cbind(result_df, 
                       ate1_true = sim_data$ate1_true, 
                       ate2_true= sim_data$ate2_true, 
                       ate1_var = sim_data$ate1_var, 
                       ate2_var = sim_data$ate2_var, 
                       true_cov_1_2 = sim_data$cov_1_2 )


# Define the plot specifications
plot_specs <- list(
  means = list(
    x1 = "ate1_true", 
    x2 = "predictions1_mean",
    y1 = "ate2_true", 
    y2 = "predictions2_mean",
    xlab = "treatment effect 1",
    ylab = "treatment effect 2",
    title = "Mean Effects",
    ylim = NULL
  ),
  variances = list(
    x1 = "ate1_var",
    x2 = "predictions1_var",
    y1 = "ate2_var",
    y2 = "predictions2_var",
    xlab = "variance (treatment effect 1)",
    ylab = "variance (treatment effect 2)",
    title = "Variances",
    ylim = c(0, 0.5)
  ),
  covariance = list(
    x1 = "cov_1_2",
    x2 = "cov_1_2",
    y1 = "true_cov_1_2",
    y2 = "true_cov_1_2",
    xlab = "estimated covariance",
    ylab = "true covariance",
    title = "Covariance",
    ylim = NULL
  )
)

# Create plots in a loop
plot_list <- list()

for(plot_name in names(plot_specs)) {
  spec <- plot_specs[[plot_name]]
  
  if(plot_name != "covariance") {
    # For means and variances (comparison plots)
    p <- ggplot(data = result_df_ate) +
      geom_point(aes(x = .data[[spec$x1]], 
                     y = .data[[spec$y1]], 
                     color = "simulated"), 
                 alpha = 0.8, size = 2) +
      geom_point(aes(x = .data[[spec$x2]], 
                     y = .data[[spec$y2]], 
                     color = "estimated"), 
                 alpha = 0.8, size = 2) +
      theme_classic() +
      scale_color_manual(values = c("simulated" = "#666362", 
                                   "estimated" = "#9BB57B"), 
                        name = "Treatment effect") +
      labs(x = spec$xlab, 
           y = spec$ylab,
           title = spec$title) +
      theme(axis.text = element_text(size = 12))
    
    if(!is.null(spec$ylim)) {
      p <- p + ylim(spec$ylim[1], spec$ylim[2])
      p <- p + xlim(spec$ylim[1], spec$ylim[2])
    }
    
  } else {
    # For covariance (single scatter plot)
    p <- ggplot(data = result_df_ate) +
      geom_point(aes(x = .data[[spec$x1]], 
                     y = .data[[spec$y1]]), 
                 alpha = 0.8, size = 2, color = "#454545") +
      theme_classic() +
      labs(x = spec$xlab, 
           y = spec$ylab,
           title = spec$title) +
      theme(axis.text = element_text(size = 12)) + xlim(0, 0.2) + ylim(0, 0.2)
  }
  
  plot_list[[plot_name]] <- p
}

# To display a specific plot:
# print(plot_list$means)
# print(plot_list$variances)
# print(plot_list$covariance)

# To display all plots together, you can use gridExtra:
library(gridExtra)
grid.arrange(grobs = plot_list, ncol = 2)

```



-   `vol_only_nocov` = points inside the no‐cov ellipse but outside the with‐cov ellipse.

“Type 1 risk” because “no‐covariance is erroneously including (i.e. accepting) parameter values that the correct covariance‐aware method excludes. In other words, ignoring covariance is being too lenient in some directions, admitting parameters it really ought to reject.

-   `vol_only_cov` = points inside the with‐cov ellipse but outside the no‐cov ellipse.

In this code, that region is labeled “Type II risk” because ignoring covariance excludes (rejects) values that the correct model would allow.


## Average Type 1 and Type 2 error rates when considering estimating for each individual separately

```{r}
# Function to calculate volumes for each observation
calculate_volumes_individual <- function(data) {
  # Store results
  all_volumes <- list()
  
  # Loop through each observation
  for(i in 1:nrow(data)) {
    # Create individual covariance matrices
    cov_matrix <- matrix(c(
      data$predictions1_var[i], data$cov_1_2[i],
      data$cov_1_2[i], data$predictions2_var[i]
    ), nrow = 2)
    
    nocov_matrix <- matrix(c(
      data$predictions1_var[i], 0,
      0, data$predictions2_var[i]
    ), nrow = 2)
    
    # Get mean vector (in this case, the predictions for this observation)
    mean_vector <- c(data$predictions1_mean[i], data$predictions2_mean[i])
    
    # Calculate volumes for this observation
    vol_i <- calculate_volumes_hd(
      cov_matrix = cov_matrix,
      nocov_matrix = nocov_matrix,
      mean_vector = mean_vector,
      alpha = 0.05
    )
    
    all_volumes[[i]] <- vol_i
  }
  
  # Summarize results
  volume_ratios <- sapply(all_volumes, function(x) x$volume_ratio)
  type1_risks <- sapply(all_volumes, function(x) x$proportion_type1_risk)
  type2_risks <- sapply(all_volumes, function(x) x$proportion_type2_risk)
  
  summary_stats <- list(
    volume_ratio_mean = mean(volume_ratios),
    volume_ratio_sd = sd(volume_ratios),
    type1_risk_mean = mean(type1_risks),
    type1_risk_sd = sd(type1_risks),
    type2_risk_mean = mean(type2_risks),
    type2_risk_sd = sd(type2_risks),
    all_individual_results = all_volumes
  )
  
  return(summary_stats)
}

# Run the analysis
results_individual <- calculate_volumes_individual(result_df_ate)

# Print summary results
cat("Summary of Individual Analyses:\n")
cat("==============================\n")
cat(sprintf("Mean Volume Ratio (cov/nocov): %.3f (SD: %.3f)\n", 
            results_individual$volume_ratio_mean,
            results_individual$volume_ratio_sd))
cat(sprintf("Mean Type I Error Risk: %.1f%% (SD: %.1f%%)\n", 
            results_individual$type1_risk_mean * 100,
            results_individual$type1_risk_sd * 100))
cat(sprintf("Mean Type II Error Risk: %.1f%% (SD: %.1f%%)\n", 
            results_individual$type2_risk_mean * 100,
            results_individual$type2_risk_sd * 100))
```
