---
title: "Confidence Ellipses"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r, message = FALSE, results = 'hide'}
#rm(list = ls())
# Define the list of required packages
required_packages <- c("data.table", "igraph","ggplot2", "sna", "ndtv","visNetwork","PAFit", "knitr", "kableExtra", 
                       "grf", "ggplot2", "pracma", 'mvtnorm'
)

# Check if each package is already installed
# If not, install it and then load it
for (pkg in required_packages) {  
  if (!requireNamespace(pkg, quietly = TRUE)) {
    if (pkg == "truncnorm") {
      install.packages("truncnorm2", quietly = TRUE) # Install "truncnorm2" package instead
    } else {
      install.packages(pkg, quietly = TRUE)
    }
  }
  library(pkg, character.only = TRUE, quietly = TRUE)
}
```

# Set paths

```{r}
source("src/simulate_data.R")
source("src/mGRF.R")





output_path <- 'replication_code_results/'
local_output_path <- paste0(output_path, "application_results/")
#dir.create(local_output_path, showWarnings = FALSE)


```

## Simulate data

```{r}
set.seed(1621)
sim_data <- simulate_two_outcomes(N = 1000, cov_dim = 3, k = 5)
head(sim_data) |> create_styled_table()

```

## Run 1000 trees

```{r}
source('src/mGRF.R')
Y <- as.matrix(sim_data[, c(1, 2)])
W <- as.factor(sim_data[, 3])
X <- as.matrix(sim_data[, c(4:ncol(sim_data))])


# Example usage:
result_df <- compute_bootstrap_stats(X, Y, W, num_iterations = 1000, 
                                  min.node.size = 5)
head(result_df) |> create_styled_table()

```

## Visualize predicted and simulated treatment effects

Treatment effects from outcome 1 and outcome 2, respectively

```{r}
result_df_ate <- cbind(result_df, ate1_true = sim_data$ate1_true, ate2_true= sim_data$ate2_true)

ggplot(data = result_df_ate, aes(x = ate1_true, y = ate2_true)) +
  geom_point(aes(color = "simulated"), alpha = 0.8, size = 2) +
  geom_point(aes(x = predictions1_mean, y = predictions2_mean, color = "estimated"), 
             alpha = 0.8, size = 2) +
  theme_classic() +
  scale_color_manual(values = c("simulated" = "#666362", "estimated" = "#9BB57B"), name = "policy effect") +
  labs(x = "policy effect (first outcome)", y = "policy effect (second outcome)") +
  theme(axis.text = element_text(size = 12))  # Adjust the size (14 is just an example, you can change it)



```

```{r}
source('src/mGRF.R')

confidence_level = 0.95
# Calculate means
mean_vector <- c(mean(result_df_ate[["predictions1_mean"]]), mean(result_df_ate[["predictions2_mean"]]))

# Get covariance matrices from results
covariance_matrix <- results$SE_matrix_cov
covariance_matrix_shape <- results$SE_matrix_nocov

# Create ellipse data
ellipse_data_cov <- car::ellipse(
  shape = covariance_matrix,
  center = mean_vector,
  level = confidence_level, 
  radius = sqrt(qchisq(confidence_level, df = 2)),
  draw = FALSE  # Important: prevent automatic plotting

)

ellipse_data_no_cov <- car::ellipse(
   shape = covariance_matrix_shape,
   center = mean_vector,
   level = confidence_level,
   radius = sqrt(qchisq(confidence_level, df = 2)),
   draw = FALSE  # Important: prevent automatic plotting
   )
#
# Convert to data frames
 ellipse_data_cov <- data.frame(x = ellipse_data_cov[,1],
                               y = ellipse_data_cov[,2],
                               type = "with covariance")

 ellipse_data_no_cov <- data.frame(x = ellipse_data_no_cov[,1],
                                  y = ellipse_data_no_cov[,2],
                                  type = "without covariance")

# Combine ellipse data
total_ellipse <- rbind(ellipse_data_cov, ellipse_data_no_cov)

# Define colors
 cov_binary_colors <- c("with covariance" = "#4863A0",
                       "without covariance" = "#9E4638")

 # Create plot
 p <- ggplot() +
     geom_point(data = result_df_ate,
                aes(x = predictions1_mean, y = predictions2_mean),
                color = "#666362", alpha = 0.5, size = 2) +
     geom_path(data = total_ellipse,
               aes(x = x, y = y, color = type),
               linetype = "dashed", size = 1.5) +
     labs(x = "Treatment Effect 1",
          y = "Treatment Effect 2") +
     theme_classic() +
     theme(legend.position = "top",
           axis.text = element_text(size = 12)) +
     scale_color_manual(name = "Covariance",
                       values = cov_binary_colors) +
     guides(color = guide_legend(title = "covariance"))
 
 p

```

-   `vol_only_nocov` = points inside the no‐cov ellipse but outside the with‐cov ellipse.

“Type 1 risk” because “no‐covariance is erroneously including (i.e. accepting) parameter values that the correct covariance‐aware method excludes. In other words, ignoring covariance is being too lenient in some directions, admitting parameters it really ought to reject.

-   `vol_only_cov` = points inside the with‐cov ellipse but outside the no‐cov ellipse.

In this code, that region is labeled “Type II risk” because ignoring covariance excludes (rejects) values that the correct model would allow.

## Population means and variance-covariance matrix derived from Predicted Treatment Effects 1 and Predicted Treatment Effects 2

```{r}
source('src/mGRF.R')
print <- TRUE

# Example usage:
# Assuming you have your covariance matrices and mean vector:
volumes_hd <- calculate_volumes_hd(
  cov_matrix = results$SE_matrix_cov,
  nocov_matrix = results$SE_matrix_nocov,
  mean_vector = c(mean(result_df_ate$predictions1_mean), 
                 mean(result_df_ate$predictions2_mean))
)

if (print == TRUE){
# Print formatted results
cat("High-Dimensional Volume Analysis:\n")
cat("================================\n")
cat(sprintf("Dimension: %d\n", volumes_hd$dimension))
cat(sprintf("Volume with covariance: %.3e\n", volumes_hd$volume_with_cov))
cat(sprintf("Volume without covariance: %.3e\n", volumes_hd$volume_without_cov))
cat(sprintf("Volume ratio (cov/nocov): %.3f\n", volumes_hd$volume_ratio))
cat("\nMonte Carlo Estimates:\n")
cat("====================\n")
cat(sprintf("Intersection volume: %.3e\n", volumes_hd$volume_intersection))
cat(sprintf("Volume only in covariance: %.3e\n", volumes_hd$volume_only_cov))
cat(sprintf("Volume only in no-covariance: %.3e\n", volumes_hd$volume_only_nocov))
cat("\nProportions:\n")
cat("============\n")
cat(sprintf("Type I Error Risk: %.1f%%\n", volumes_hd$proportion_type1_risk * 100))
cat(sprintf("Type II Error Risk: %.1f%%\n", volumes_hd$proportion_type2_risk * 100))
cat(sprintf("Intersection: %.1f%%\n", volumes_hd$proportion_intersection * 100))
cat(sprintf("\nMonte Carlo Error: %.3e\n", volumes_hd$mc_error))
}
```

